---
title: "hw1_yx2857"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# load libraries
library(tidyverse)
library(ggplot2)
```


```{r data}
# import dataset
data("penguins", package = "palmerpenguins")
```

**Problem 1**   

The dataset, _penguins_, stores information of penguins, including species, island, body measurements, sex, and year. The _penguin_ dataset has `r ncol(penguins)` columns and `r nrow(penguins)` rows.  
The mean flipper length is `r sprintf("%.2f", mean(penguins$flipper_length_mm, na.rm = T))` mm.  
Another dataset, _penguins_raw_, stores the raw data that _penguins_ originates from, including more detailed information of the penguins. The _penguins_raw_ dataset has `r ncol(penguins_raw)` columns and `r nrow(penguins_raw)` rows.  


```{r plot}
# Make a scatterplot of flipper_length_mm (y) vs bill_length_mm (x); color points using the species variable

p <- ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species))
p <- p + geom_point()
p <- p + labs(title = "Flipper Length vs. Bill Length", x = "Bill Length (mm)", y = "Flipper Length (mm)", color = "Species")

# save the plot
ggsave("flipper_bill_length.png", plot = p, width = 6, height = 4)

print(p)
```

**Problem 2**  
Dataset "store" is created as follows:  
```{r problem2_data}
# Create a data frame comprised of:
# 
# a random sample of size 10 from a standard Normal distribution
# a logical vector indicating whether elements of the sample are greater than 0
# a character vector of length 10
# a factor vector of length 10, with 3 different factor “levels”

# set a random seed
set.seed(8105)

balance = rnorm(10)
good = balance > 0
staff = letters[1:10]
manager = factor(c(rep(letters[1:3], 3), letters[1]))

store <- tibble(balance, good, staff, manager)

store

```

Try to take the mean of each variable in your dataframe. What works and what doesn’t?  
```{r problem2_mean}
# Try to take the mean of each variable in your dataframe. What works and what doesn’t?
# 
# Hint: to take the mean of a variable in a dataframe, you need to pull the variable out of the dataframe. Try loading the tidyverse and using the pull function.

print("Mean of variable, 'balance'")
mean(pull(store, balance), na.rm = T)
print("Mean of variable, 'good'")
mean(pull(store, good), na.rm = T)
print("Mean of variable, 'staff'")
mean(pull(store, staff), na.rm = T)
print("Mean of variable, 'manager'")
mean(pull(store, manager), na.rm = T)
```

Taking the mean works for the numeric and logical variables, `balance` and `good`, but not for the variable `staff` or `manager`, which are character and factor variables respectively.  

```{r problem2_convert, include=TRUE, echo = TRUE, results=FALSE}
# Write a code chunk that applies the as.numeric function to the logical, character, and factor variables (please show this chunk but not the output). What happens, and why? Does this help explain what happens when you try to take the mean?

as.numeric(pull(store, balance))
as.numeric(pull(store, good))
as.numeric(pull(store, staff))
as.numeric(pull(store,manager))
```

When applying the `as.numeric` function to the logical, character, and factor variables, the logical variable `good` is converted to 0 and 1, the character variable `staff` is converted to `NA`, and the factor variable `manager` is converted to the integer representation of the factor levels. This helps explain what happens when trying to take the mean of these variables. The `mean` function only works for numeric and logical variables, and it will not work for character or factor variables.

